Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-04-02 GWLDCI.tex
added bibliography to both articles

This paper is the first in a two part series investigating the combinatorics and geometry underlying SYM $N=4$ theory. This series lays out several results about the relationship between the Wilson loop diagrams and the positroid cells that give a CW complex structure to the positive Grassmanian. This paper expands on an observation from \cite{wilsonloops}, that shows that certain families of Wilson loop diagrams correspond to the same positroid cell. This paper gives a way of enumerating the number of different Wilson loop diagrams that correspond to a given positroid cell. Furthermore, it shows that that the number of cells in the image of the map to positive Grassmannians is given by counting the directions of associahedrons.

The second paper in this series identifies which positroid cell is associated to which Wilson loop diagram \cite{GWLDCII}. It proves that a Wilson loop diagram with $k$ propagators corresponds to a $3k$ dimensional positroid cell. Furthermore, it gives an algorithm for passing from a Wilson loop diagram to the Grassman necklace that identifies the associated positroid cell. In this manner the authors give a means of identifying which positroid cell is associate to which Wilson loop diagram that circumvents the painstaking process of examining bases sets. Finally, the authors characterize the the volume forms that Wilson loop diagrams associate to each positroid cells.

In recent years, there has been a an active program researching the geometry and combinatorics underlying SYM $N=4$ theory \cite{wilsonloops, amplituhedron, correlahedron, karpwilliams, galashinlam}. This body of work started with the observation that the on shell amplitudes of this theory correspond to the volume of a subspace of a positive Grasmannian, called an Amplituhedron \cite{amplituhedron}. Since then, there has been significant work to study the structure of the amplituhedron, geometrically and combinatorially, for instance \cite{amplituhedron2, bigamplituhedronpaper, karpwilliams, galashinlam}. The focus of much of this work has been to understand the relationship between the combinatorial objects from the physics of the system, BCFW diagrams, and their corresponding natural object for studying the positive Grassmannians, positroid cells.

Meanwhile, there has been a different body of work studying SYM $N=4$ theory from the point of view of Feynman integrals in twistor space\cite{BullimoreSkinner, AldayMaldecena?, Wittenpaper}. These integrals are calculated via Wilson loops. The first author and Amat uncovered a connection between the Feynman integrals developed in this literature and positroids, defined  as matroids that can be realized as an element of $\Gr(k,n)$ \cite{wilsonloops}. With the second author and others, this work was extended to study these Feynman integrals in terms of positroid cells and the positive Grassmanians \cite{casestudy, non-orientable}. Others have tried different approaches to define a geometric space associated to these integrals in a manner similar to the Amplituhedron \cite{correlahedron, HeslopStewart}.

Both the Amplituhedron literature and the Wilson loop literature associate $n$ particle, $N^kMHV$ interaction in SYM $N=4$ theory to volume forms on a set of positroid cells of $\Gr(k,n)$. This gives a CW complex of some submanifold in $\Gr(k,n)$. Understanding the geometry and topology of this space is the subject of ongoing research. For Wilson loop diagrams with $k=2$ and $n=6$, one knows that the manifold is not contractible \cite{casestudy}. Furthermore, the spaces defined by the Amplituhedron and the Wilson loop diagrams are not the same \cite{correlahedron}. In particular, it it more natural to consider Deodhar cells of $\mathbb{G}(k, n+1)$ for Wilson loop diagrams. In this case, one gets that the manifold represented by the diagrams is not orientable \cite{non-orientable}.

In this paper, we focus on the correspondence between Wilson loop diagrams and positroids, first defined in \cite{wilsonloops}. In that paper, the authors show that there are certain families of diagrams, specifically those with exact subdiagrams, that do map to the same positroid cell. In \cite{casestudy} the authors explictly show that not every positroid cell of a given dimension corresponds to a Wilson loop diagram. In this paper, we better enumerate this correspondence. We show that the map from Wilson loop diagrams to postroid cells is one to one exactly on the subset of diagrams with no non-trivial exact subdiagrams. We show that exact diagrams correspond to uniform matroids, and that contraction by exact subdiagrams has special matroidal properties. Further, given a Wilson loop diagram with a non-trivial exact subdiagram, we count the total number of Wilson loop diagrams that correspond to that same positroid cell. While we do not count the number of distint positroid cells that correspond to Wilson loop diagrams, we do show that this enumeration is equivalent to enumerating the number of directions of an associahedron.

@@

\begin{thebibliography}{10}

\bibitem{wilsonloop}
S. Agarwala, E. M. Amat.
\newblock Wilson loop diagrams and positroids.
\newblock {\em Commun. Math. Phys.} 2016.

\bibitem{reversingOh}
S. Agarwala, S. Fryer.
\newblock An algorithm to construct the Le diagram associated to a Grassmann necklace
\newblock \url{arXiv:1803.01726}

@@

\bibitem{Stanley}
  R. P. Stanley,
  \newblock Enumerative Combinatorics, Vol. 2
  \newblock Wadsworth (1999)

\bibitem{Ziegler}
  G. M. Ziegler,
  \newblock Lectures on Polytopes
  \newblock Springer GTM 152, (1995)

\bibitem{CSZinequivalent}
  Cesar Ceballos, Francisco Santos, and G\"unter M. Ziegler,
  \newblock Many non-equivalent realizations of the associahedron
  \newblock {\em Combinatorica}, 35(5):513--551, (2015).  \url{arxiv:1109.5544}

>>>>

This paper is the first in a two part series investigating the combinatorics and geometry underlying SYM $N=4$ theory. This series lays out several results about the relationship between the Wilson loop diagrams and the positroid cells that give a CW complex structure to the positive Grassmanian. This paper expands on an observation from \cite{wilsonloop}, that shows that certain families of Wilson loop diagrams correspond to the same positroid cell. This paper gives a way of enumerating the number of different Wilson loop diagrams that correspond to a given positroid cell. Furthermore, it shows that that the number of cells in the image of the map to positive Grassmannians is given by counting the directions of associahedrons.

The second paper in this series identifies which positroid cell is associated to which Wilson loop diagram \cite{generalcombinatoricsI}. It proves that a Wilson loop diagram with $k$ propagators corresponds to a $3k$ dimensional positroid cell. Furthermore, it gives an algorithm for passing from a Wilson loop diagram to the Grassman necklace that identifies the associated positroid cell. In this manner the authors give a means of identifying which positroid cell is associate to which Wilson loop diagram that circumvents the painstaking process of examining bases sets. Finally, the authors characterize the the volume forms that Wilson loop diagrams associate to each positroid cells.

In recent years, there has been a an active program researching the geometry and combinatorics underlying SYM $N=4$ theory \cite{wilsonloop, Arkani-Hamed:2013jha, Amplituhedronsquared, AmplituhedronDecomposition, galashinlam}. This body of work started with the observation that the on shell amplitudes of this theory correspond to the volume of a subspace of a positive Grasmannian, called an Amplituhedron \cite{Arkani-Hamed:2013jha}. Since then, there has been significant work to study the structure of the amplituhedron, geometrically and combinatorially, for instance \cite{Arkani-Hamed:2013kca, arkani:2012nw, AmplituhedronDecomposition, galashinlam}. The focus of much of this work has been to understand the relationship between the combinatorial objects from the physics of the system, BCFW diagrams, and their corresponding natural object for studying the positive Grassmannians, positroid cells.

Meanwhile, there has been a different body of work studying SYM $N=4$ theory from the point of view of Feynman integrals in twistor space\cite{Adamo:2011pv, Britto:2005fq, Cachazo:2004kj}. These integrals are calculated via Wilson loops. The first author and Amat uncovered a connection between the Feynman integrals developed in this literature and positroids, defined  as matroids that can be realized as an element of $\Gr(k,n)$ \cite{wilsonloop}. With the second author and others, this work was extended to study these Feynman integrals in terms of positroid cells and the positive Grassmanians \cite{casestudy, non-orientable}. Others have tried different approaches to define a geometric space associated to these integrals in a manner similar to the Amplituhedron \cite{Amplituhedronsquared, HeslopStewart}.

Both the Amplituhedron literature and the Wilson loop literature associate $n$ particle, $N^kMHV$ interaction in SYM $N=4$ theory to volume forms on a set of positroid cells of $\Gr(k,n)$. This gives a CW complex of some submanifold in $\Gr(k,n)$. Understanding the geometry and topology of this space is the subject of ongoing research. For Wilson loop diagrams with $k=2$ and $n=6$, one knows that the manifold is not contractible \cite{casestudy}. Furthermore, the spaces defined by the Amplituhedron and the Wilson loop diagrams are not the same \cite{Amplituhedronsquared}. In particular, it it more natural to consider Deodhar cells of $\mathbb{G}(k, n+1)$ for Wilson loop diagrams. In this case, one gets that the manifold represented by the diagrams is not orientable \cite{non-orientable}.

In this paper, we focus on the correspondence between Wilson loop diagrams and positroids, first defined in \cite{wilsonloop}. In that paper, the authors show that there are certain families of diagrams, specifically those with exact subdiagrams, that do map to the same positroid cell. In \cite{casestudy} the authors explictly show that not every positroid cell of a given dimension corresponds to a Wilson loop diagram. In this paper, we better enumerate this correspondence. We show that the map from Wilson loop diagrams to postroid cells is one to one exactly on the subset of diagrams with no non-trivial exact subdiagrams. We show that exact diagrams correspond to uniform matroids, and that contraction by exact subdiagrams has special matroidal properties. Further, given a Wilson loop diagram with a non-trivial exact subdiagram, we count the total number of Wilson loop diagrams that correspond to that same positroid cell. While we do not count the number of distint positroid cells that correspond to Wilson loop diagrams, we do show that this enumeration is equivalent to enumerating the number of directions of an associahedron.

@@

%\begin{thebibliography}{10}
%\bibliographystyle{plain}
%\bibliography{C:/Users/agarws1/Documents/GitHub/OldWLDPapers/Bibliography}

\begin{thebibliography}{50}

\bibitem{Adamo:2011pv}
Tim Adamo, Mathew Bullimore, Lionel Mason, and David Skinner.
\newblock {Scattering Amplitudes and Wilson Loops in Twistor Space}.
\newblock {\em J.Phys.A}, A44:454008, 2011.

\bibitem{wilsonloop}
Susama Agarwala and Eloi~Marin Amat.
\newblock Wilson loop diagrams and positroids.
\newblock {\em Communications in Mathematical Physics}, 2016.
\newblock arXiv:1509.06150.

\bibitem{casestudy}
Susama Agarwala and Sian Fryer.
\newblock A study in $\textrm{Gr}_{\geq 0}(2,6)$: from the geometric case book
  of {W}ilson loop diagrams and {SYM} $n=4$.
\newblock arXiv:1803.00958.

\bibitem{generalcombinatoricsI}
Susama Agarwala, Sian Fryer, and Karen Yeats.
\newblock Combinatorics and geometry of wilson loop diagrams.
\newblock manuscript in progress.

\bibitem{non-orientable}
Susama Agarwala and Cameron Marcott.
\newblock Wilson loops in sym $n=4$ do not parametrize an orientable space.
\newblock arXiv:1807.05397.

\bibitem{arkani:2012nw}
Nima Arkani-Hamed, Jacob~L. Bourjaily, Freddy Cachazo, Alexander~B Goncharov,
  Alexander Postnikov, and Jaroslav Trnka.
\newblock Scattering amplitudes and the positive {G}rassmannian.
\newblock arXiv:1212.5605, 2012.

\bibitem{Arkani-Hamed:2013kca}
Nima Arkani-Hamed and Jaroslav Trnka.
\newblock {Into the Amplituhedron}.
\newblock {\em JHEP}, 12:182, 2014.

\bibitem{Arkani-Hamed:2013jha}
Nima Arkani-Hamed and Jaroslav Trnka.
\newblock {The Amplituhedron}.
\newblock {\em JHEP}, 10:30, 2014.

\bibitem{Britto:2005fq}
Ruth Britto, Freddy Cachazo, Bo~Feng, and Edward Witten.
\newblock {Direct proof of tree-level recursion relation in Yang-Mills theory}.
\newblock {\em Phys.Rev.Lett.}, 94:181602, 2005.

\bibitem{Cachazo:2004kj}
Freddy Cachazo, Peter Svrcek, and Edward Witten.
\newblock {MHV vertices and tree amplitudes in gauge theory}.
\newblock {\em JHEP}, 0409:006, 2004.

\bibitem{Amplituhedronsquared}
Burkhard Eden, Paul Heslop, and Lionel Mason.
\newblock The correlahedron.
\newblock 2017.
\newblock arxiv:1701.00453.

\bibitem{galashinlam}
P.~{Galashin} and T.~{Lam}.
\newblock {Parity duality for the amplituhedron}.
\newblock {\em arXiv e-prints}, May 2018.

\bibitem{HeslopStewart}
Paul Heslop and Alastair Stewart.
\newblock The twistor wilson loop and the amplituhedron.
\newblock {\em Journal of High Energy Physics}, 2018(10):142, Oct 2018.

\bibitem{AmplituhedronDecomposition}
Steven~N. Karp, Lauren~K. Williams, and Yan~X. Zhang.
\newblock Decompostions of amplituhedra.
\newblock arXiv:1708.09525.

@@

\bibitem{Stanley}
  R. P. Stanley.
  \newblock Enumerative Combinatorics, Vol. 2
  \newblock Wadsworth (1999)

\bibitem{Ziegler}
  G. M. Ziegler.
  \newblock Lectures on Polytopes
  \newblock Springer GTM 152, (1995)

\bibitem{CSZinequivalent}
  Cesar Ceballos, Francisco Santos, and G\"unter M. Ziegler.
  \newblock Many non-equivalent realizations of the associahedron
  \newblock {\em Combinatorica}, 35(5):513--551, (2015).  \url{arxiv:1109.5544}

### 2019-04-02 GWLDCII.tex
added bibliography to both articles

In recent years, there has been a some interest in understanding the geometry and combinatorics underlying the field theory SYM $N=4$ \cite{wilsonloops, amplituhedron, correlahedron, karpwilliams, galashinlam}. This started with the observation BCFW diagrams, which represent the on shell interactions of this theory correspond to plabic graphs \cite{bigamplituhedronpaper.} Plabic graphs are one of the many equivalent ways to identify positroid cells of the positie Grassmannians \cite{Postnikov}. This work led to the representation of the on shell integrals associated this theory in terms of a geometric space embedded in $\Gr(k,n)$, called the Amplituhedron \cite{amplituhedron, amplituhedron2}. More over, inegrals associated to the BCFW diagrams define a volume form on the Amplituhedron Since then, there has been significant work to understand the volume form and its physical implications \cite{galashinlam, others}.

Meanwhile, a different body of work has studied SYM $N=4$ theory from the point of view of Feynman integrals in twistor space \cite{BullimoreSkinner, AldayMaldecena?, Wittenpaper}. These integrals are calculated via holomorphic Wilson loops. As with the Amplituhedron, each Wilson loop represents the sum of a family of $N^kMHV$ Feynman diagrams. The sum of the corresponding sum of Feynman integrals associate a volume form to space \cite{Hodges?}. The first author and Amat uncovered a connection between the Feynman integrals developed in this literature and positroids, defined  as matroids that can be realized as an element of $\Gr(k,n)$ \cite{wilsonloops}, while other work has shown a connection between the space parametrized by Wilson loop diagrams and the Amplituhedron \cite{casestudy, correlahedron}.

In the above work and elsewhere, it is clear that while the space parametrized by the Wilson loop diagrams is related to the Amplituhedron. They are not the same. For in stance, \cite{non-orientable} shows that the space parametrized by the Wilson loop diagrams can be seen as a non-orientable vector bundle over some submanifold of $\Gr(k,n)$, while it is conjectured that the Amplituhedron is orientable. Furthermore, the cells associated to the Wilson loop diagrams are different from the cells associated to BCFW diagrams, as one sees from counting arguments in comparing the results of \cite{KarpWiliams} and \cite{casestudy}. This is not too alarming, as one expects space parametrized by the Wilson loop diagrams to project down to something close to the Amplituhedron \cite{wilsonloops, correlahedron}.

@@

\begin{thebibliography}{10}

\bibitem{wilsonloop}
S. Agarwala, E. M. Amat.
\newblock Wilson loop diagrams and positroids.
\newblock {\em Commun. Math. Phys.} 2016.

>>>>

In recent years, there has been a some interest in understanding the geometry and combinatorics underlying the field theory SYM $N=4$ \cite{wilsonloop, Arkani-Hamed:2013jha, Amplituhedronsquared, AmplituhedronDecomposition, galashinlam}. This started with the observation BCFW diagrams, which represent the on shell interactions of this theory correspond to plabic graphs \cite{GrassmannAmplitudebook}. Plabic graphs are one of the many equivalent ways to identify positroid cells of the positie Grassmannians \cite{Postnikov}. This work led to the representation of the on shell integrals associated this theory in terms of a geometric space embedded in $\Gr(k,n)$, called the Amplituhedron \cite{Arkani-Hamed:2013jha, Arkani-Hamed:2013kca}. More over, inegrals associated to the BCFW diagrams define a volume form on the Amplituhedron. Since then, there has been significant work to understand the volume form and its physical implications \cite{galashinlam, UnwindingAmplituhedron}.

Meanwhile, a different body of work has studied SYM $N=4$ theory from the point of view of Feynman integrals in twistor space \cite{Adamo:2012xe, Adamo:2011pv,  Boels:2007qn, Bullimore:2010pj}. These integrals are calculated via holomorphic Wilson loops. As with the Amplituhedron, each Wilson loop represents the sum of a family of $N^kMHV$ Feynman diagrams. The sum of the corresponding sum of Feynman integrals associate a volume form to space \cite{Hodges:2012ym}. The first author and Amat uncovered a connection between the Feynman integrals developed in this literature and positroids, defined  as matroids that can be realized as an element of $\Gr(k,n)$ \cite{wilsonloop}, while other work has shown a connection between the space parametrized by Wilson loop diagrams and the Amplituhedron \cite{casestudy, Amplituhedronsquared}.

In the above work and elsewhere, it is clear that while the space parametrized by the Wilson loop diagrams is related to the Amplituhedron. They are not the same. For in stance, \cite{non-orientable} shows that the space parametrized by the Wilson loop diagrams can be seen as a non-orientable vector bundle over some submanifold of $\Gr(k,n)$, while it is conjectured that the Amplituhedron is orientable. Furthermore, the cells associated to the Wilson loop diagrams are different from the cells associated to BCFW diagrams, as one sees from counting arguments in comparing the results of \cite{AmplituhedronDecomposition} and \cite{casestudy}. This is not too alarming, as one expects space parametrized by the Wilson loop diagrams to project down to something close to the Amplituhedron \cite{wilsonloop, Amplituhedronsquared}.

@@

%\bibliographystyle{plain}
%\bibliography{C:/Users/agarws1/Documents/GitHub/OldWLDPapers/Bibliography}

\begin{thebibliography}{10}
\bibitem{Adamo:2011pv}
Tim Adamo, Mathew Bullimore, Lionel Mason, and David Skinner.
\newblock {Scattering Amplitudes and Wilson Loops in Twistor Space}.
\newblock {\em J.Phys.A}, A44:454008, 2011.

\bibitem{Adamo:2012xe}
Tim Adamo and Lionel Mason.
\newblock {Twistor-strings and gravity tree amplitudes}.
\newblock {\em Class.Quant.Grav.}, 30:075020, 2013.

\bibitem{wilsonloop}
Susama Agarwala and Eloi~Marin Amat.
\newblock Wilson loop diagrams and positroids.
\newblock {\em Communications in Mathematical Physics}, 2016.
\newblock arXiv:1509.06150.

\bibitem{casestudy}
Susama Agarwala and Sian Fryer.
\newblock A study in $\textrm{Gr}_{\geq 0}(2,6)$: from the geometric case book
  of {W}ilson loop diagrams and {SYM} $n=4$.
\newblock arXiv:1803.00958.

\bibitem{non-orientable}
Susama Agarwala and Cameron Marcott.
\newblock Wilson loops in sym $n=4$ do not parametrize an orientable space.
\newblock arXiv:1807.05397.

\bibitem{UnwindingAmplituhedron}
Nima Arkani-Hamed, Hugh Thomas, and Jaroslav Trnka.
\newblock Unwinding the amplituhedron in binary.
\newblock {\em Journal of High Energy Physics}, 2018(1):16, Jan 2018.
\newblock arXiv:1704.05069.

\bibitem{Arkani-Hamed:2013kca}
Nima Arkani-Hamed and Jaroslav Trnka.
\newblock {Into the Amplituhedron}.
\newblock {\em JHEP}, 12:182, 2014.

\bibitem{Arkani-Hamed:2013jha}
Nima Arkani-Hamed and Jaroslav Trnka.
\newblock {The Amplituhedron}.
\newblock {\em JHEP}, 10:30, 2014.

\bibitem{Boels:2007qn}
Rutger Boels, Lionel Mason, and David Skinner.
\newblock {From twistor actions to MHV diagrams}.
\newblock {\em Phys.Lett.}, B648:90--96, 2007.

\bibitem{Bullimore:2010pj}
Mathew Bullimore, Lionel Mason, and David Skinner.
\newblock {MHV Diagrams in Momentum Twistor Space}.
\newblock {\em JHEP}, 1012:032, 2010.

\bibitem{Amplituhedronsquared}
Burkhard Eden, Paul Heslop, and Lionel Mason.
\newblock The correlahedron.
\newblock 2017.
\newblock arxiv:1701.00453.

\bibitem{galashinlam}
P.~{Galashin} and T.~{Lam}.
\newblock {Parity duality for the amplituhedron}.
\newblock {\em arXiv e-prints}, May 2018.

\bibitem{HeslopStewart}
Paul Heslop and Alastair Stewart.
\newblock The twistor wilson loop and the amplituhedron.
\newblock {\em Journal of High Energy Physics}, 2018(10):142, Oct 2018.

\bibitem{Hodges:2012ym}
Andrew Hodges.
\newblock {A simple formula for gravitational MHV amplitudes}.
\newblock 2012.

\bibitem{AmplituhedronDecomposition}
Steven~N. Karp, Lauren~K. Williams, and Yan~X. Zhang.
\newblock Decompostions of amplituhedra.
\newblock arXiv:1708.09525.

\bibitem{GrassmannAmplitudebook}
Freddy Cachazo Alexander Goncharov Alexander Postnikov Jaroslav~Trnka
  Nima~{A}rkani {H}amed, Jacob~Bourjaily.
\newblock {\em Grassmannian Geometry of Scattering Amplitudes}.
\newblock Cambridge University Press, Cambridge, UK, 2016.
